US Korea Exchange Rates \\ STA206 Final Project

### 2014-12-13 paper.tex
report

We started out with some general summary statistics and plotting. The first
issue was missing data, which we had anticipated. While most of the
variables were reported monthly, \texttt{debt} was reported on a quarterly
basis and 
Our simple approach for dealing with this was to
make a plot of these variables over time.
If they are approximately linear then we could 
linearly interpolate with respect to time. The scaled plots 
\ref{fig:na_plot_KOR} and
\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
In this plots one can observe that $debt_USA$ is decreasing; this is
because $debt$ carries a sign. The value of $debt_USA$ on

When beginning this project we knew that the time dependency and
multicollinearity would be the
biggest obstacles, since in this case the assumption that the errors
$\epsilon_i \sim N(0, \sigma^2)$ does not hold.
Figure~\ref{fig:scatterplot} shows what the original data looks like.
We observe clear patterns emerging, with some even approaching a
continuous mathematical functional relationship- ie the patterns
between \texttt{gdp}, \texttt{gdp\_deflator}, and \texttt{debt} for Korea.
Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
correlations for the original data set- they are

Our first approach for overcoming this time dependence was to create a new
data frame from the original data by taking the difference between each
subsequent row. More formally, if our original design matrix $X$ is $n
\times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
setting
\[
    \tilde{X_{i,*}} = X_{(i + 1), *} - X_{i, *}
\]
for $i = 1, 2, \dots, n-1$ where $X_{i, *}$ represents the $i$th row of
$X$.

This approach did indeed eliminate the most obvious patterns, but 
sacrifices interpretability

Our first step was to separate the data into training and validation sets,
and only

Figure~\ref{fig:na_plot_KOR} shows the variables which were interpolated.

@@

@@

@@

>>>>

We started out with some general summary statistics and plotting. The first
issue was missing data, which we had anticipated. While most of the
variables were reported monthly, \texttt{debt} was reported on a quarterly
basis and 
Our simple approach for dealing with this was to
make a plot of these variables over time.
If they are approximately linear then we could 
linearly interpolate with respect to time. The scaled plots 
Figure~\ref{fig:na_plot_KOR} and
Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
because \texttt{debt} carries a sign. The value of \texttt{debt\_USA} on

When beginning this project we knew that the time dependency and
multicollinearity would be the
biggest obstacles, since in this case the assumption that the errors
$\epsilon_i \sim N(0, \sigma^2)$ does not hold.
Figure~\ref{fig:scatterplot} shows what the original data looks like.
We observe clear patterns emerging, with some even approaching a
continuous mathematical functional relationship- ie the patterns
between \texttt{gdp}, \texttt{gdp\_deflator}, and \texttt{debt} for Korea.
Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
correlations for the original data set- they are highly correlated. In
fact, 30 \% of the columns have absolute sample correlation greater than 0.9.

Our first approach for overcoming this time dependence was to create a new
data frame from the original data by taking the difference between each
subsequent row (monthly difference). 
More formally, if our original design matrix $X$ is $n
\times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
setting
\[
    \tilde{X_{i,*}} = X_{(i + 1), *} - X_{i, *}
\]
for $i = 1, 2, \dots, n-1$ where $X_{i, *}$ represents the $i$th row of
$X$.

As can be observed in Figure~\ref{fig:diff_scatterplot},
this approach did indeed eliminate the most obvious patterns, but 
at the expense of interpretability. We attempted regression on this
transformed data, but were unsuccessful.

\subsection{Ratios}

Because exchange rate is a ratio of Korean currency / US dollar we decided
to make transformed variables consisting of the Korean economic indicator
divided by the corresponding US indicator. We also included
\texttt{inflation} untransformed after some preliminary analysis 
indicated that it was
important. \texttt{Date} was intentionally left out of the model in order
to be able to interpret the model in terms of the other parametrs. However,
the effects of time are still present through the variables that are linear
over time.

\subsection{Analysis}

Our first step when fitting the model was to separate the data into
training and validation sets. We checked to make sure that the variables
looked approximately the same in both training and validation sets.

We used forward stepwise selection to choose a model. 
Based on the criteria from the output of the \texttt{leaps regsubsets}
in the appendix we chose to explore a
model with an intercept and 5 predictor variables. The one with 5
predictors had the lowest BIC of all the models, and the Mallow's CP at
5.39 was significantly better than the model with 4 predictors which had a
CP value of 13.91. Since 5.39 is close to 5 and 13.91 is pretty far from 4
we conclude that the model with 5 variables is correct, meaning that it has
little bias.

The variables under further investigation are:
\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
\texttt{inflation\_USA}, and \texttt{govspending}.

The box - cox transformation in figure~\ref{fig:boxcox} suggests a
transformation of $1 / x^2$, but this extreme and is due to the outliers.
We used the multiplicative inverse transformation and examined the
residual plots figure~\ref{fig:cooks} to see why this is the case.

We see
that the points 113 and 117 are the primary outlying cases. Looking up the
corresponding dates reveals that these are the points occuring during
October 2008 and February 2009, during the subprime mortgage crisis
which shook the global economy. There are no points between 113 and 117
because the test train split didn't select them. In the validation set we
discovered similar results for this period. The subprime mortgage crisis is
the reason that they are outliers.

@@

Our use of dynamic data caused trouble because the data was a little too
dynamic.

@@

Model Selection Output:
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & sse & bic & r2 & r2a & cp \\
  \hline
1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
  2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
  3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
  4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
  5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
  6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
  7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
  8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
  9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
  10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
   \hline
\end{tabular}
\end{table}

@@

\begin{figure}
  \centering
    \includegraphics{diff_scatterplot.pdf}
  \caption{Scatterplot of data after row difference transformation}
  \label{fig:diff_scatterplot}
\end{figure}

\begin{figure}
  \centering
    \includegraphics{boxcox.pdf}
  \caption{Box cox procedure on training set- affected by outliers}
  \label{fig:boxcox}
\end{figure}

\begin{figure}
  \centering
    \includegraphics{cooks.pdf}
  \caption{Cooks distance for residuals of training data}
  \label{fig:cooks}
\end{figure}
